---
title: "Class11"
author: "Ryan Fong"
date: "2022-11-02"
output: pdf_document
---

# Section 1. Proportion on G/G in a population

Download a CSV file from Ensemble < https://uswest.ensembl.org/Homo_sapiens/Variation/Sample?db=core;r=17:39894595-39895595;v=rs8067378;vdb=variation;vf=105535077;sample=mxl#373531_tablePanel >

Here we read this CSV file

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand)
```

```{r}
table(mxl$Genotype..forward.strand) / nrow(mxl)
```

Now let's look at a different population. I picked the GBR.

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")

```

Find proportion of G|G 

```{r}
table(gbr$Genotype..forward.strand.) /nrow(gbr)
```

The variant that is associated with childhood asthma is more frequent in GBR population than in MXL population.

## Section 4: Population Scale Analysis

>Q13: Read this file into R and determine the sample size for each genotype and their
corresponding median expression levels for each of these genotypes. 


How many samples do we have?

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head(expr)
```
We want to check total genotype

```{r}
nrow(expr)
```
See how many of each type

```{r}
table(expr$geno)
```

ggplot2 wil first be loaded
```{r}
library(ggplot2)

```


Now a boxplot will be made

```{r}
ggplot(expr) + aes(x=geno,exp, fill=geno) +
  geom_boxplot(notch=TRUE)
```

From the boxplot, the median expression levels for the A/A genotype is roughly 32. For the A/G genotype, it is 25. For the G/G genotype it is 20

>Q14: Generate a boxplot with a box per genotype, what could you infer from the relative
expression value between A/A and G/G displayed in this plot? Does the SNP effect the
expression of ORMDL3?

I can infer from this relative expression value between A/A and G/G in this plot is that A/A has more expression when compared to G/G. SNP does have an effect on the expression of ORMDL3 because there are increases and decreases in expression in other genese from SNP.










