---
title: "Class10"
author: "Ryan Fong"
format: pdf
---

We will first download and import the candy data
```{r}
candy_file <- "candy-data.csv"
candy = read.csv(candy_file, row.names=1)
head(candy)
```

>Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

>Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

The winpercent variable can be used to see a popularity of a candy relative to the other candies.
```{r}
candy["Twix", ]$winpercent
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

My favorite candy are air heads.
```{r}
candy["Air Heads", ]$winpercent
```

>Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

The skimr package will be downloaded so the `skimr()` function can be used to find the overview of the dataset.

```{r}
library("skimr")
skim(candy)
```


>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, it would be winpercent.

>Q7. What do you think a zero and one represent for the candy$chocolate column?

0 represents that there are no chocolate in the candy and 1 means that there are chocolate in the candy.

>Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent) +
  geom_histogram(bins=10, col="red",fill="orange")
```

>Q9. Is the distribution of winpercent values symmetrical?

Not really because the distribution is slightly skewed to the right.

>Q10. Is the center of the distribution above or below 50%?

The center of distribution is below 50 percent

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate.inds <- as.logical(candy$chocolate)
chocolate.wins <- candy[chocolate.inds,]$winpercent
mean(chocolate.wins)
```

```{r}
fruity.inds <- as.logical(candy$fruity)
fruity.wins <- candy[fruity.inds,]$winpercent
mean(fruity.wins)
```

Chocolate candy is higher ranked than fruity candy

>Q12. Is this difference statistically significant?

```{r}
t.test(chocolate.wins,fruity.wins)
```

Yes, because the p-value is small.

>Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```

>Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```

>Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col()
```

Now colors will be added
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "green"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
#my_cols
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```

>Q17. What is the worst ranked chocolate candy?

sixlets

>Q18. What is the best ranked fruity candy?

starbursts

A plot is made to compare the winpercent and pricepercent to compare the best value candy. First, ggrepel must be downloaded

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 25)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Resses Miniatures

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```

In the top 5 most expensive candies listed above, Nik L Nip is the least popular

corrplot will be use to see how the variables are related to one another. corrplot will first be downloaded
```{r}
library(corrplot)
## corrplot 0.90 loaded
cij <- cor(candy)
corrplot(cij)
```

>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

chocolate and fruity

>Q23. Similarly, what two variables are most positively correlated?

chocolate and winpercent

PCA will be applied using `prcomp()`
```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

A plot will be made
```{r}
plot(pca$x[,1],pca$x[,2])
```

Character changed and colors added
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

A new data frame is made so seperate columns can be included to make ggplot look nicer
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
```

Now this can be ploted setting ggplot equal to p
```{r}
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

ggrepel in implemented to add the labels of the candy on the plot
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

plotly is used to generate an interactive which will be downloaded first
```{r}
library(plotly)
ggplotly(p)
```

A barplot is made to compared the PCA combination of each category
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```

>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, hard, and pluribus. This makes sernse because these 3 categories have little to no correlation to the other categories in PC1 by the negative direction.











